---
title: "Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
is_ghactions <- identical(Sys.getenv("CI"), "true")

```

## Install

Install the R package `dwc.ar4gw` by running the following code:

```{r install_and_load, eval = FALSE}
### Optionally: specify GitHub Personal Access Token (GITHUB_PAT)
### See here why this might be important for you:
### https://kwb-r.github.io/kwb.pkgbuild/articles/install.html#set-your-github_pat

# Sys.setenv(GITHUB_PAT = "mysecret_access_token")

# Install package "remotes" from CRAN
if (! require("remotes")) {
  install.packages("remotes", repos = "https://cloud.r-project.org")
}

# Install KWB package 'kwb-r/dwc.ar4gw' from GitHub
remotes::install_github("kwb-r/dwc.ar4gw", dependencies = TRUE)

```

## Define paths

```{r define_paths}
# Load the R package
library(dwc.ar4gw)

### define paths
paths_list <- list(
  tdir = tempdir(), 
  cloud_dir = "dwc/Work-Packages/WP3_Public_Involvement/AR4GW/ModelData/MF2005",
  local_dir = "C:/kwb/projects/<cloud_dir>",
  download_dir = tempdir(),
  modflow_dir = "<root_data>",
  model_name = "MB16",
  mfout = "<modflow_dir>/<model_name>",
  mf_cbc = "<mfout>.cbc",
  mf_nam = "<mfout>.nam",
  mf_bas = "<mfout>.bas",
  mf_dis = "<mfout>.dis",
  text_heads = "<mfout>.fhd",
  text_drawdown = "<mfout>.fdn",
  binary_cell.by.cell = "<mfout>.cbc",
  binary_heads = "<mfout>.bhd",
  binary_drawdown = "<mfout>.bdn"
)

paths <- kwb.utils::resolve(paths_list, root_data = paths_list$download_dir)
paths_local <- kwb.utils::resolve(paths_list, root_data = paths_list$local_dir)
```

## Modflow Files

This tutorial needs a few `MODFLOW` files which can be provided [locally](#locally)
from a predefined folder on your Windows computer or directly downloaded from
the DWC [Cloud](#cloud). Within this tutorial the latter approach is used. 

Currently all input files in the MODFLOW `.nam` file are required and two output 
files:

* `<model_name>`.`nam` (input: name file)

* `<model_name>`.`dis` (input: discretisation file)

* `<model_name>`.`lst`|`bas`|`oc`|`gmg`|`bcf`|`wel`|`riv`|`rch`|`evt` (not loaded, 
but need to be available in model directory)

* `<model_name>`.`cbc` (output: cell by cell budget fie)

* `<model_name>`.`bhd` (output: binary heads file)



where for `<model_name>` we use `r paths_list$model_name`, which is defined in 
the `paths_list` in the previous chapter [Define Paths](#define-paths).


<!-- According to [ModelMuse documentation]( -->
<!-- https://water.usgs.gov/nrp/gwsoftware/ModelMuse/Help/index.html?head_and_drawdown_panes.htm) -->
<!-- `heads` and `drawdowns` will be saved with the following `file extensions`:  -->

<!-- * **heads**: `.fhd` if it is a formatted text file and `.bhd` if it is a binary file.  -->

<!-- * **drawdowns**: `.fdn` if it is a formatted text file and `.bdn` if it is a binary file. -->

### Locally

Copy the content of the DWC cloud folder `r paths$cloud_dir` to 
`r paths_local$local_dir` as shown below:

```{r modflow_local, echo = FALSE}
paths_local
```

```{r modflow_local_rename, eval = FALSE}
paths <- paths_local
```

### Cloud 

You can download the required model files from the DWC cloud if you are a 
registered user with access to the folder ``r paths$cloud_dir``

For doing so follow the steps below: 

1. Open `RStudio` and run `usethis::edit_r_environ()`

2. In the opened window add the required environment variables

```
NEXTCLOUD_URL = "https://<replace-with-dwc-cloud-url>"
NEXTCLOUD_USER = "<your-dwc-cloud-username>" # your username
NEXTCLOUD_PASSWORD = "your-nextcloud-app-password" ### see details below
```

For creating `<your-nextcloud-app-password>`:

* go to: https://replace-with-dwc-cloud-url/index.php/settings/user/security

* scroll down to `create new app password`

* select a name e.g. `r-script` and copy the token and replace `your-nextcloud-app-password`


3. Finally you need to restart Rstudio and proceed with the code below:

```{r nextcloud_download}

required_mffiles <- paste0(paste0(c("nam", "dis", "lst", "bas", "oc", "gmg",
         "bcf", "wel", "riv", "rch", "evt", "bhd", "cbc"), 
       collapse = "$|"), "$")


# Download .cbc and .bhd and .dis files
mf_files <- kwb.nextcloud::list_files(
  paths$cloud_dir,
  full_info = TRUE) %>%
  dplyr::filter(stringr::str_detect(.data$file,
                                    pattern = sprintf("^%s\\.",  
                                                      paths$model_name))) %>% 
  dplyr::filter(stringr::str_detect(.data$file,
                                    pattern = required_mffiles))

mf_files

kwb.nextcloud::download_files(href = mf_files$href,
                              target_dir = paths$download_dir)

```


## Flopy release

### Install flopy

Install latest `flopy` release from `conda` [https://anaconda.org/conda-forge/flopy](https://anaconda.org/conda-forge/flopy)-

```{r install_flopy_ghactions, eval = is_ghactions}

### create an environment "flopy" for installing "flopy" 
kwb.python::conda_py_install(env_name = "flopy", 
                             pkgs = list(conda = c("flopy", "python", "pyshp"), 
                                         py = NULL))

```

### Use flopy

```{r use_flopy, eval = is_ghactions}
#reticulate::use_miniconda("flopy", required = TRUE)
# reticulate::py_help(object = flopy$utils$postprocessing$get_extended_budget)
extended_budget <- dwc.ar4gw::get_extended_budget(cbcfile = paths$binary_cell.by.cell) 
summary(extended_budget$time_1.0$Qx_ext)

str(extended_budget)

flopy <- dwc.ar4gw::import_flopy()

# reticulate::py_help(object = flopy$utils$binaryfile$HeadFile)
heads <- flopy$utils$binaryfile$HeadFile(filename = paths$binary_heads,
                                         verbose = TRUE)

# reticulate::py_help(object = heads$get_times)
times <- heads$get_times()[[1L]]

dat_heads <- stats::setNames(
  lapply(times, function(time) {
  heads$get_data(totim = time) }),
  nm = sprintf("time_%.1f", times)
  )

str(dat_heads)
summary(dat_heads$time_365.0)
summary(dat_heads$time_365.0)[c(1,6)]
# reticulate::py_help(object = flopy$modflow$mf$Modflow$load)
mf <- flopy$modflow$mf$Modflow$load(f = paths$mf_nam, 
                                    model_ws = dirname(paths$mf_nam),
                                    load_only = "dis", #c("dis", "bas6", "bcf6"),
                                    forgive = TRUE,
                                    verbose = TRUE
                                    )



# reticulate::py_help(object = flopy$utils$postprocessing$get_gradients)
gradients <- flopy$utils$postprocessing$get_gradients(
  heads = dat_heads$time_365.0,
  m = mf,
  nodata = c(-1e+30,1e+30))

str(gradients)
summary(gradients)

```

### Export 

Model to shapefile (gets really big !!!!)

```{r export, eval = FALSE}
mf$export(f = "MB16.shp")
```

For details on the used `flopy` functions have a look at the documentation here:

[https://flopy.readthedocs.io/en/latest/source/flopy.utils.postprocessing.html](https://flopy.readthedocs.io/en/latest/source/flopy.utils.postprocessing.html)


## Flopy development 

Install  `flopy development` version from GitHub [https://github.com/modflowpy/flopy](https://github.com/modflowpy/flopy)

### Install flopy-dev



```{r install_flopy_dev_install_ghactions, eval = FALSE}

### create an environment "flopy" for installing "flopy" 
kwb.python::conda_py_install(env_name = "flopy_dev", 
                             pkgs = list(conda = c("python", "pyshp"), 
                                         py = "https://github.com/modflowpy/flopy/zipball/develop/801a41705d8979c6982fb8c2955d56225d971218"))

```
### Use flopy-dev

```{r install_flopy_dev_ghactions, eval = FALSE}

reticulate::use_miniconda(condaenv = "flopy_dev", required = TRUE)
flopy_dev <- reticulate::import("flopy", convert = TRUE)


kstpkper <- c(0,0)

# reticulate::py_help(object = flopy_dev$utils$postprocessing$get_extended_budget)
budget_array <- flopy_dev$utils$postprocessing$get_extended_budget(
      cbcfile = paths$binary_cell.by.cell, kstpkper = kstpkper) 

# reticulate::py_help(object = flopy_dev$utils$binaryfile$HeadFile)
heads <- flopy_dev$utils$binaryfile$HeadFile(filename = paths$binary_heads,
                                         verbose = TRUE)

# reticulate::py_help(object = heads$get_data)
heads_array <- heads$get_data(kstpkper = kstpkper)

# reticulate::py_help(object = flopy_dev$modflow$mf$Modflow$load)
mf <- flopy_dev$modflow$mf$Modflow$load(f = paths$mf_nam, 
                                    model_ws = dirname(paths$mf_nam),
                                    load_only = c("dis", "bas6", "bcf6"),
                                    forgive = TRUE,
                                    verbose = TRUE
                                    )

 
# reticulate::py_help(object = flopy$utils$postprocessing$get_specific_discharge)
# Works now with flopy 3.3.4 release candidate: 
# https://github.com/modflowpy/flopy/blob/b6092cbc94039750de309a832c7f554751782ac1/flopy/utils/postprocessing.py#L606
# reticulate::py_help(object = flopy_dev$utils$postprocessing$get_specific_discharge)
specific_discharge <- setNames(flopy_dev$utils$postprocessing$get_specific_discharge(
  vectors = budget_array, 
  model = mf, 
  head = heads_array
  ), 
  nm = c("qx", "qy", "qz")
)
# stats for all layers
summary(specific_discharge$qx)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -3837       0       0       0       0    1340  742928 
summary(specific_discharge$qy)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -3967       0       0       0       0    1744  743864 
summary(specific_discharge$qz)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#   -0.6     0.0     0.0     0.0     0.0     0.2  737240

#top layer
layer <- 1
summary(as.vector(specific_discharge$qx[layer,,])) 
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#   -0.6     0.0     0.0     0.0     0.0     0.2  737240
summary(as.vector(specific_discharge$qy[layer,,])) 
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -1.55   -0.03    0.00   -0.01    0.01    1.84   92983 
summary(as.vector(specific_discharge$qz[layer,,])) 
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -0.30    0.00    0.00    0.00    0.00    0.09   92155

# layer2
layer <- 2
summary(as.vector(specific_discharge$qx[layer,,])) 
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -0.23    0.00    0.00    0.00    0.00    0.28   92866
summary(as.vector(specific_discharge$qy[layer,,])) 
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -0.23    0.00    0.00    0.00    0.00    0.19   92983
summary(as.vector(specific_discharge$qz[layer,,]))
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -0.55    0.00    0.00    0.00    0.00    0.17   92155 


### Calculate flow velocity 
# Set effective porosity for each layer
# layer 1: 0.3 (unconfined)
# layer 2-8: 0.0001 (confined)
effective_porosity <- c(0.3, rep(0.0001, 7))
n_layers <- mf$nlay

flow_velocity <- setNames(lapply(specific_discharge, function(q) {
  for (layer in seq_len(n_layers)) {
    q[layer,,] <- q[layer,,] / effective_porosity[layer]
  }
  q
}), nm = c("vx", "vy", "vz")
)

# stats for all layers
summary(flow_velocity$vx)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max.      NA's 
#-38372483       -24         0       -29         8  13398450    742928 
summary(flow_velocity$vy)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max.      NA's 
#-39667686       -24         0       -44        11  17443782    743864 
summary(flow_velocity$vz)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#-5858.8    -1.2    -0.2    -1.5     0.0  1749.5  737240 

#top layer
layer <- 1
summary(as.vector(flow_velocity$vx[layer,,])) 
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -5.44   -0.06   -0.01   -0.02    0.01    5.89   92866
summary(as.vector(flow_velocity$vy[layer,,])) 
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -5.16   -0.08   -0.01   -0.03    0.02    6.13   92983 
summary(as.vector(flow_velocity$vz[layer,,])) 
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#  -1.01    0.00    0.00    0.00    0.00    0.29   92155

# layer2
layer <- 2
summary(as.vector(flow_velocity$vx[layer,,])) 
#    Min.  1st Qu.   Median     Mean  3rd Qu.     Max.     NA's 
#-2288.47   -42.81    -0.34   -23.69     0.42  2842.85    92866 
summary(as.vector(flow_velocity$vy[layer,,])) 
#    Min.  1st Qu.   Median     Mean  3rd Qu.     Max.     NA's 
#-2337.17   -43.87    -0.36   -42.00     1.08  1852.84    92983 
summary(as.vector(flow_velocity$vz[layer,,]))
#    Min.  1st Qu.   Median     Mean  3rd Qu.     Max.     NA's 
#-5533.12    -3.01    -1.62    -4.79    -0.66  1749.48    92155


### Print as tidy data.frame 
flow_velocity_vx_long <- dwc.ar4gw::to_long(flow_velocity$vx)
head(flow_velocity_vx_long)
#  layer col row value
#1     1   1   1   NaN
#2     2   1   1   NaN
#3     3   1   1   NaN
#4     4   1   1   NaN
#5     5   1   1   NaN
#6     6   1   1   NaN

flow_velocity_vx_wide <- dwc.ar4gw::to_wide(flow_velocity_vx_long, 
                                            parameter = "vx")
flow_velocity_vx_wide
## A tibble: 534,640 x 10
#   column   row  vx_1  vx_2  vx_3  vx_4  vx_5  vx_6  vx_7  vx_8
#    <int> <int> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
# 1      1     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
# 2      2     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
# 3      3     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
# 4      4     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
# 5      5     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
# 6      6     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
# 7      7     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
# 8      8     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
# 9      9     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
#10     10     1   NaN   NaN   NaN   NaN   NaN   NaN   NaN   NaN
## ... with 534,630 more rows

### Plot data of all layers
dwc.ar4gw::plot_data(heads_array, title = "Heads")
dwc.ar4gw::plot_data(flow_velocity$vx, title = "Flow velocity x")
dwc.ar4gw::plot_data(flow_velocity$vy, title = "Flow velocity y")
dwc.ar4gw::plot_data(flow_velocity$vz, title = "Flow velocity z")

```
